### 2022-12-28 paper.Rmd
minor

Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide). High between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets, that should however be as similar as possible with regard to response-relevant attributes [@lahl2006]. Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data. That is, individual stimuli are numerically coded on relevant dimensions [@dry2009]. For stimuli in psycholinguistic experiments, for example, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982]. In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977]. In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019]. Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli. A different anticlusting approach relies on clustering objectives that can directly be computed using the attribute values. K-means is probably the best-known objective of this kind. In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic [@jain2010; @steinley2006; @brusco2006branch]. Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np]. Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful approach when aiming for high between-group similarity [@spath1986; @valev1998; @valev1983]. Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters. However, other distribution characteristics---such as the variance---are not targeted. @papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually resulted in more similar variances. When requiring overall between-group similarity, the negligence of all distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].

@@

# Plot the results, by N
ldf %>% 
  group_by(method, Objective, N) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>%
  ggplot(aes(x = N, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16)

>>>>

Oftentimes, partitioning applications in psychology are focused on ensuring between-group similarity rather than within-group heterogeneity (even though both goals oftentimes coincide). High between-group similarity is for example desirable when preparing experimental conditions that use different stimulus sets, that should however be as similar as possible with regard to response-relevant attributes [@lahl2006]. Rather than relying on pairwise (dis)similarity ratings, stimulus selection tasks are usually conducted on the basis of attribute data. That is, individual stimuli are numerically coded on relevant dimensions [@dry2009]. For stimuli in psycholinguistic experiments, attributes may consist of ratings for imagery and concreteness, as well as orthographic variables [@friendly1982]. In other cases, attributes are binary and may represent the presence or absence of a feature [@tversky1977]. In the context of anticlustering, one approach for handling attribute data is to compute an appropriate dissimilarity measure such as the pairwise Euclidean or squared Euclidean distance across the set of attributes [@brusco2019]. Subsequently, a distance-based anticlustering criterion such as the diversity can be optimized to obtain a partitioning of the stimuli. A different anticlusting approach relies on clustering objectives that can directly be computed using the attribute values. K-means is probably the best-known objective of this kind. In k-means clustering, the sum of the squared Euclidean distances between data points and their cluster centers is minimized, usually using the k-means heuristic [@jain2010; @steinley2006; @brusco2006branch]. Minimizing the k-means criterion simultaneously maximizes between-cluster separation and within-cluster homogeneity [@aloise2009np]. Reversing the k-means objective function---using maximization instead of minimization---has been recognized as a useful approach when aiming for high between-group similarity [@spath1986; @valev1998; @valev1983]. Specifically, as @spath1986 noted, k-means anticlustering minimizes differences with regard to the means of the numeric attributes across clusters. However, other distribution characteristics---such as the variance---are not targeted. @papenberg2020 showed that k-means anticlustering tends to over-optimize similarity with regard to the mean; even an entirely random group assignment usually resulted in more similar variances. When requiring overall between-group similarity, the negligence of all distribution characteristics other than the mean is misguided, because similar means can be obtained even when the underlying distributions are clearly different [e.g., @anscombe1973].

@@

# Plot the results, by N
axis_ticks <- seq(24, 300, by = 12)
axis_labels <- axis_ticks
axis_labels[c(FALSE, TRUE)] <- ""

ldf %>% 
  group_by(method, Objective, N) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>%
  ggplot(aes(x = N, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(
    breaks = axis_ticks, 
    labels = axis_labels
  )
